Experiences with Planning for Natural Language Generation

### 2009-06-03 trunk/Papers/ci-spark/experiments.tex
changes

In the second experiment, we evaluate the performance of the planners on
problems arising in the GIVE domain. We construct a series of test worlds,
similar to the one illustrated in Figure~\ref{fig:give-maps}(a). These
worlds consist of a $2n$ by $h$ grid of positions, such that there are
buttons at positions $(2i-1,1)$ and $(2i,h)$ for $1 \leq i \leq n$. The
player starts in position $(1,1)$ and must press all the buttons
to successfully complete the game. The world is generated as a GIVE world
description, and then automatically converted into a planning problem by
the GIVE software. For instance, Figure~\ref{fig:give-planning} shows some
of the actions available in the GIVE domain in PDDL syntax.

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are shown
in Figure~\ref{fig:give-minimal}. The most obvious result is that
FF is unable to solve any problems beyond $n=13$ on our experimentation
machine within the memory limit of 1 GB. SGPLAN, on the other hand, solves
instances beyond $n=40$ without major problems. In this case, our
implementation of GraphPlan is unable to find any of these plans. The time
spent on grounding is not a major factor in either planner, probably
because the planners need more time to actually compute the plan---for
instance, the optimal plan for the problem $n=40$ has a length of about
1600 steps. As a concrete example, the following is a minimal plan for the
case of a $2$ by $2$ grid with $2$ buttons for the player to press (i.e.,
$n=1$, $h=2$):

@@

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal.pdf}}
  \caption{Results for the unordered and ordered minimal GIVE domains with grid
  height $h=20$. The horizontal axis is the grid width, $b = 2n$. The
  vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal}
\end{figure}

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal-grounding.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal-grounding.pdf}}
  \caption{Comparison of the total running time and grounding time for selected
  planners in the $h=20$ minimal GIVE domain. The horizontal axis is the grid
  width, $b = 2n$. The vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal-grounding}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.85\columnwidth]{graph-exp4}
  %\includegraphics[width=1\columnwidth]{pic-runtime-empty-world}
  \caption{Results on the GIVE domain with inaccessible
    positions for $h=20$ and $n=5$. The horizontal axis is $w$.
    The vertical axis is the runtime in seconds.}
  \label{fig:give-runtime-nosoln}
\end{figure}

>>>>

In the second set of experiments, we evaluate the performance of several
planners on problems arising in the GIVE domain. Besides the FF, Metric-FF, and
SGPLAN planners we used in the previous set of experiments, we expand our
testing to include FF(h$_a$), LAMA, and C$^3$ \todo{citations}. Each of these
additional planners competed in the deterministic ``sequential, satifying''
track of the 2008 International Planning Competition; all planners generally
performed well on the competition domains, with LAMA the overall winner of
the track.\footnote{See \texttt{http://ipc.informatik.uni-freiburg.de/} for
details of the 2008 IPC.}

In the first GIVE test, we construct a series of grid worlds, similar to the one
illustrated in Figure~\ref{fig:give-maps}(a). These worlds consist of a $2n$ by
$h$ grid of positions, such that there are buttons at positions $(2i-1,1)$ and
$(2i,h)$ for $1 \leq i \leq n$. The player starts in position $(1,1)$ and must
press all the buttons to successfully complete the game.  (The actions in this
domain are similar to the PDDL actions in Figure~\ref{fig:give-planning}.)
We consider two variants of this problem in our tests. In the \emph{unordered}
problem, the player is permitted to press the buttons in any order to
successfully achieve the goal. In the \emph{ordered} version of the problem, the
player is unable to initially move to any grid cell containing a button, except
for the cell containing the first button, $u_1$. Pressing $u_1$ releases the
position of the next button, $l_1$, allowing the player to move into this cell.
Similarly, pressing button $l_1$ frees button $u_2$, and so on. The end result
is a set of constraints that forces the buttons to be pressed in a particular
order to achieve the goal. As a concrete example, the following is a minimal
plan (in either variant of the problem) for the case of a $2$ by $2$ grid with
$2$ buttons (i.e., $n=1$, $h=2$):

@@

Results for the $h=20$ case, with the grid width $N$ ranging from $0$ to $40$,
are shown in Figure~\ref{fig:give-minimal}. In the unordered case
(Figure~\ref{fig-give-minimal}(a)), the most obvious result is that some of the
planners tested---Metric-FF, FF(h$_a$), and C$^3$---are unable to solve any
problems beyond $N=24$ on our experimentation machine within the memory limit of
2 GB.\footnote{\todo{describe testing environment}} While FF, Lama, and SGPLAN
are all able to solve problem instances up to $N=40$, the total running time
varies greatly between this planners. For instance, FF takes almost 35
seconds to solve the $N=40$ problem, LAMA takes around 6.5 seconds, and SGPLAN
shows impressive performance, taking well less than a second.

In a real NLG system, where response time is

machine within the memory limit of 1 GB. SGPLAN, on the other hand, solves
instances beyond $n=40$ without major problems. In this case, our
implementation of GraphPlan is unable to find any of these plans. The time
spent on grounding is not a major factor in either planner, probably
because the planners need more time to actually compute the plan---for
instance, the optimal plan for the problem $n=40$ has a length of about
1600 steps.

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal.pdf}}
  \caption{Results for the unordered and ordered minimal GIVE domains with grid
  height $h=20$. The horizontal axis is the grid width, $N = 2n$. The
  vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal}
\end{figure}

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal-grounding.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal-grounding.pdf}}
  \caption{Comparison of the total running time and grounding time for selected
  planners in the $h=20$ minimal GIVE domain. The horizontal axis is the grid
  width, $N = 2n$. The vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal-grounding}
\end{figure}

@@

%\begin{figure}
%  \centering
%  \includegraphics[width=0.85\columnwidth]{graph-exp4}
%  %\includegraphics[width=1\columnwidth]{pic-runtime-empty-world}
%  \caption{Results on the GIVE domain with inaccessible
%    positions for $h=20$ and $n=5$. The horizontal axis is $w$.
%    The vertical axis is the runtime in seconds.}
%  \label{fig:give-runtime-nosoln}
%\end{figure}
